Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-27 demystify.tex
updates

\pagestyle{empty}
\begin{tikzpicture}[overlay,remember picture]

@@

          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).

@@

Assuming you're not familiar with any particular one, then first search for profs. that share similar research interests.  For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.  If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.  Then you research about that prof. (e.g., go to their website) and contact them (\autoref{sec:contact}).

@@

  In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
  \tcblower
  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities. Many will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
\end{commentbox}

@@

\subsubsection{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.  The main difference is their focus: research, teaching, or both. Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students. Thus when contacting a prof. for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).

@@

\section{Contacting a prof. How to get a desired reply?}\label{sec:contact}

\subsectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}

@@

Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).

@@

\subsubsection*{Additional Resources}
\begin{itemize}
  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
\end{itemize}

@@

    \item \textbf{Self-focus}. Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}). Mention why you're interested in their work and how your background can contribute.

    \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded.\\
    Personally I do not want to see course transcript or test scores in the first email. If I am interested I'll ask for them. Attaching CV is OK. Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant to my work.  
    \item \textbf{Flowery greetings and language}. Don't use ``Dear esteemed professor'', don't ``hey there''.  Do not call the prof. by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).  Do not use Mr., Mrs., etc. To be safe, use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).

@@

\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.  You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).

@@

You can also take honor thesis or independent study courses with a professor.  This is a good way to get research experience and also get a LoR from the professor.  You can also ask your academic adviser or other faculty members for suggestions.  Finally, you can also ask your peers who are already doing research.  Use the method described in \autoref{sec:contact} to contact professors

>>>>

%\pagestyle{empty}

\begin{tikzpicture}[overlay,remember picture]

@@

          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{chap:contact}).

@@

Assuming you're not familiar with any particular one, then first search for profs. that share similar research interests.  For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.  If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.  Then you research about that prof. (e.g., go to their website) and contact them (\autoref{chap:contact}).

@@

  In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
  \tcblower
  \textbf{Vu}: \autoref{chap:contact} shows how to contact a professor for research opportunities. Many will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
\end{commentbox}

@@

\subsubsection{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.  The main difference is their focus: research, teaching, or both. Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students. Thus when contacting a prof. for research opportunities (\autoref{chap:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).

@@

\chapter{Contacting a prof. How to get a desired reply?}\label{chap:contact}

\chapterinfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}

@@

Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
See~\autoref{sec:kiss-of-death-emails} for common mistakes in emails and~\autoref{sec:interpreting-response} for interpreting replies.

@@

\subsection*{Additional Resources}
\begin{itemize}
  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
\end{itemize}

@@

    \item \textbf{Self-focus}. Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{chap:contact}). Mention why you're interested in their work and how your background can contribute.

    \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded. Don't attach course transcript or test scores in the first email. If they are  interested they will ask for them.  Attaching CV is OK. Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant.

    \item \textbf{Flowery greetings and language}. Don't use ``Dear esteemed professor'', don't ``hey there''.  Do not call the prof. by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).  Do not use Mr., Mrs., etc. To be safe, use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).

@@

\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.  You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{chap:contact}).

@@

You can also take honor thesis or independent study courses with a professor.  This is a good way to get research experience and also get a LoR from the professor.  You can also ask your academic adviser or other faculty members for suggestions.  Finally, you can also ask your peers who are already doing research.  Use the method described in \autoref{chap:contact} to contact professors
